Automatic Matching of Bullet Lands

### 2016-04-20 revision/response.Rmd

4. Random forests, as used and described on pages 18-21 of the paper comes out looking like a winner, but other than that success---worth reporting *somewhere*---, there is no new information on that method. No new insights. And certainly no comparisons with the tens of other things that could have been tried.

Reviewer B:

5. The authors should discuss their work in the context of Nicholas Petraco’s work as well as other ongoing research endeavors in impression and pattern evidence. See https://www.ncjrs.gov/pdffiles1/nij/grants/239048.pdf The alignment method should have been compared with a standard method- See Patraco’s NIJ report for a couple of examples. As presented the novel component of the paper is the ``twicing" for grooves combined with a loess fit to the remove curves from bullet images. I am not sure this is the authors intention. It would be interesting to compare this method with some of the classical methods from functional data analysis used for curve registration.

***Using the cross correlation function as a method of alignment is actually pretty standard - the approriate references are now included.***

@@

***Done for figure 15 and 16***

Reviewer 1:

21. First, the term "land" is used in the title but not defined until p. 3.  I recommend changing the title, to ensure broad comprehenson; perhaps "Automatic Matching of Bullet Marks," or something of that kind.

***To clarify the title, we have modified the abstract to provide a brief definition of bullet lands.***

22. p. 3, l. -10:  Be explicit that grooves and striations are synonyms.

***Grooves and striations are actually not synonyms in this case. The grooves are impressions induced by the rifling of the gun barrel, and properties associated with them are class characteristics not unique to individual gun barrels. We've taken care to ensure the distinction is clear between the striae of interest, which are extracted from the lands of the bullet (between grooves), and the grooves themselves.***

23. Early on in the paper, it might be helpful to mention the typical width of a striation, the number of striations typically seen on a fired bullet, and so forth. 
***EH: Add typical width of groove + lands***

@@

***EH: Presumably the typical width is some kind of class characteristic dependent on the gun itself, right?***
***HH: I am not sure about the width of a striation - the width depends on where it is measured (in both x and z direction). I'm sure there are definitions of that available. We might need some more literature.***

24. p. 2, l. -1:  It would be good to reword so the sentence doesn't start with "x3p".  Also p. 3, l. 2.

***DONE***

25. p. 3, Fig. 1:  The numbers on the y-axis are superimposed.

***We have removed the values from the y-axis, as they are not necessary in interpreting the image.***

@@

30. p. 13, l. -3:  You write "Given that striation patterns are typically much larger, ..." and I think it would be more clear to write "Given that striation widths are typically much larger, ...".

***DONE***

31. p. 16, l. -1/-2:  Could you elaborate on why a "large number of maximal CMS by itself is not indicative of a match"?  I think the reason is that there could be other marks that are completely incompatible and preclude a match, based on the next sentence, but I am not sure and a few words would be helpful.

***In the Hamby study, with 10,212 land-to-land comparisons which are known non-matches, two of these comparisons still yielded a CMS of 12. Figure 13 emphasizes this. In our particular data, all comparisons with a CMS of 13 or more were in fact matches. Still, it is likely due to pure chance alone that, particularly if more bullets were used in comparison, a small portion would have high pairwise CMS despite being non-matches.***

### 2016-04-20 imaging-paper.Rnw
Update response, make more paper changes

\address{Address of the Authors\\
Department of Statistics and Statistical Laboratory\\ Snedecor Hall\\
Ames Iowa 50011-1210\\
\printead{e1}\\
\phantom{E-mail:\ }\printead*{e2}\\
\phantom{E-mail:\ }\printead*{e3}}
\end{aug}
%
\begin{abstract}
In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination. Firearm examination is a forensic tool used to help the court determine whether two bullets were fired from the same gun barrel. During the firing process, rifling, manufacturing defects, and impurities in the barrel create striation marks on the bullet. Identifying these striation markings in an attempt to match two bullets is one of the primary goals. We propose an automated framework for the analysis of the 3D surface measurements, that first transcribes the markings into a 2D plotting framework. This makes identification of matches easier and allows for a quantification of both matches and matchability of barrels. The automatic matching routine we propose manages to (a) correctly identify lands with too much damage to be suitable for comparison, and (b) correctly identify all 10,384 land-to-land matches of the James Hamby study \citep{hamby:2009}.
\end{abstract}

>>>>

\address{Address of the Authors\\
Department of Statistics and Statistical Laboratory\\ Snedecor Hall\\
Ames Iowa 50011-1210\\
\printead{e1}\\
\phantom{E-mail:\ }\printead*{e2}\\
\phantom{E-mail:\ }\printead*{e3}}
\end{aug}
%
\begin{abstract}
In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination. Firearm examination is a forensic tool used to help the court determine whether two bullets were fired from the same gun barrel. During the firing process, rifling, manufacturing defects, and impurities in the barrel create striation marks on the bullet. Identifying these striation markings in an attempt to match two bullets is one of the primary goals. We propose an automated framework for the analysis of the 3D surface measurements, that first transcribes the markings into a 2D plotting framework. This makes identification of matches easier and allows for a quantification of both matches and matchability of barrels. The automatic matching routine we propose manages to (a) correctly identify lands (the surface between two bullet grooves) with too much damage to be suitable for comparison, and (b) correctly identify all 10,384 land-to-land matches of the James Hamby study \citep{hamby:2009}.
\end{abstract}
